A sensor grid approach for large scale acoustic monitoring

### 2017-07-18 paper/article.tex
ml proof read starts and conclusion

\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
\newcommand{\fg}[1]{\textcolor{red}{ Felix: #1}}

@@

The acoustic environment represents both societal and environmental concerns. Monitoring the acoustic environnment in urban but also rural or wilderness areas is therefore an important matter. Fostering on the recent development of low cost hardware acoustic sensors, we propose to consider a sensor grid approach to tackle this issue.

In this kind of approach, the crucial question is the nature of the data that is transmitted from the sensors to the processing and archival servers. To this end, we propose an efficient audio coding scheme based on third octave band spectral representation that allows 1) the estimation of most acousitc indicators and 2) the recognition of acoustic events at state of the art performance rate. The former is useful to provide quantitative information about the acoustic environment, while the latter is useful to gather more qualitative information and build perceptually motivated indicators using for example some emergence of a given sound source.

The coding scheme is also demonstrated to transmit spectrally encoded data that, reverted to the time domain using state of the art techniques, is not intelligible, thus protecting the privacy of citizens.

@@

The advent of low cost acoustic sensors together with the need to better monitor and comprehend the acoustic environment of urban and wilderness areas give rise to the deployment of experimental sensor networks such as the sonyc (wp.nyu.edu/sonyc) and cense (cense.ifsttar.fr) projects.

@@

In our application setting, the requirements are the following. First,  the number of sensor nodes shall be extended as needed without having to change the hardware and software architecture, \textit{i.e.} the approach shall be scalable. Second, the nodes shall be energy efficient in order ot ease the deployment of the sensors grid to the desired topology, ideally autonomous. Third, the encoding scheme used to transmit the data from the sensor to the storage and processing servers shall be designed with care in order to ensure the privacy of the citizen. Lastly, the data stored shall be as rich as possible while remaining a low bitrate for efficient transmission and storage.

@@

The latter scheme has several benefits. First, the sensor is much simpler and can thus be autonomous in terms of energy, easing the deployment of the network. Second, it allows researchers to gather large amount of data that can be post processed and studied further offline. Data can be re analysed following newer classification schemes or using new indicators.

However, transmitting the raw audio through the network has several disavantages, in terms of required bandwidth, storage capabilities and privacy. Thus, for transmission from the sensors to the storage unit, the data can be encoded in a more efficient way than lossy audio encoding standards \cite{pan1995tutorial} as the audio signal is not mandatory for the computation of the acoustic indicators and the features needed for event recognition. Also, as the data is transmitted using the network and stored, one must ensure that the intelligibility of potential speech utterances is lost during the coding process, in order to ensure the privacy of the citizens.

We thus propose in this paper a audio coding scheme specifically tailored for use in a sensor grid which has the following features. It is of low bitrate but still allows the computation of most of the standard acoustics indicators with high precision. As far as acoustic event detection is concerned, we report equivalent performance of several state of the art classification schemes from features computed using raw audio data and encoded data. Finally, according to preliminary perceptual evaluation, the proposed coding scheme very strongly degrades the intelligibility, thus ensuring citizen privacy.

In order to promote reproducible research, the coder as well as the experiments needed to generate the figures will be made available to the community. The remaining of the paper is organized as follows: Section \ref{sec:background} reviews the types of data that needs to be estimated using the transmitted data and Section \ref{sec:coder} describes the proposed encoding scheme. The experimental protocol designed to validate the proposed approach is described in Section \ref{sec:protocol} and results are discussed in Section \ref{sec:results}.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figures/censeCoder}
\vspace{-0.15in}
\caption{Alternative implementations of the sensor grid approach for the monitoring of the acoustic environnment. In A) the raw audio is transmitted, in B) only the detected events, and in C) a compressed spectral representation. Thickness of arrows indicate bandwidth and thickness of boxes indicates the level of computation or storage required.}
\label{fig:codingScheme}
\end{figure}

@@

The present work is constrained to allow both acoustic monitoring and sound event recognition in urban soundscapes. Here we briefly review state-of-the-art methods in both subjects in order to better motivate our choices.

@@

The considered sensor network primarily aims at monitoring urban soundscapes, that is, continously assessing their content and impact on the population. In the literature, this is typically enabled by measuring energetic acoustic indicators such as the equivalent sound pressure level $L_{eq}$ in dB SPL or its A-weighted equivalent $L_{Aeq}$ in dBA. However, while these indicators have proven to correlate well with perceptual evaluations for negative impact sound environments \cite{gozalo2015}, they are not sufficient to fully describe soundscapes \cite{rychtarikova2013}. Many other variables can be derived to better account for previously implicit properties \cite{can2016} including percentile values or time evolution. Studies have been conducted to select relevant subsets of descriptors in sound environment characterization \cite{can2015, brocolini2013, nilsson2007}.\\

@@

Eventhough the above discussed indicators provide more richness than $L_{eq}$, recent studies show that abstract statistics are still not sufficient to fully model the human perception of soundscapes. The recognition of sources of interest is thus of importance. As far as the encoding scheme is concerned, it is thus important that the encoded data allows the computation of the above cited indicators but also the recognition of sources of interest with state of the art methods.

The recognition of sources of interest from audio streams has been the subject of extensive research in the past on speech \cite{anusuya2009}, music \cite{tzanetakis2002}, and lately more complex scenes in which the current work falls. Studied classification methods are diverse, ranging from time-dependant modeling with HMMs \cite{ntalampiras2014} to "bag-of-frames" approaches \cite{aucouturier2007, foggia2015}. Common architectures include learning-based classifiers such as support vector machines (SVM) \cite{kumar2016}, gaussian mixture models (GMM) \cite{radhakrishnan2005} or neural networks \cite{salamon2017, piczak2015}. However, the selection of relevant features is still an open debate. The most used are certainly spectral \cite{khunarsal2013} or cepstral \cite{couvreur2004} representations of the signal. Among them, mel spectrograms and their cepstral-domain derivation, the mel-frequency cepstrum coefficients (MFCC), are the most recurrent. These representations effectively try to model the human cochlear response to sounds by grouping frequency components around critical bands in a logarithmic scale, thus bearing important physical significance. They may also be exploited together with features computed in other domains to better model signal properties. For instance, \cite{cai2006} adds spectral features related to harmonicity and salient frequencies, and \cite{chu2009} uses a matching pursuit (MP) algorithm to deduce time-domaid features. Another promising solution is feature engineering via unsupervised learning, which \cite{salamon2015-2} implements with a k-means clustering technique. Alternative data representations such as the scattering transform \cite{bauge2013} show good results in environmental sound classification tasks \cite{salamon2015}.\\
A more detailed review of used methods is available in \cite{chachada2013}.

@@

The technical constraints of size inherent to the studied setup make it impossible to transmit raw audio recordings. Following the considerations briefly exposed in the previous section, we conclude that third-octave bands are a reasonable choice. As mentionned, they provide advantages in both acoustic monitoring quality and reduced data volume per measurement period. The physical content is also close to that of mel spectrograms, being just another filterbank transform with logarithmic frequency scaling.\\

@@

The first step is to represent the signal in the frequency domain using a short-term Fourier transform (STFT). The continuously sampled audio is segmented into 125~ms frames to provide "fast" acoustic measurements. The data is zero-padded to the next power of two to allow best fft performances. No overlap is used as it doubles both computational and memory costs and is not required for our application. A windowing function can be applied, for instance a rectangular window to ensure energy conservation in a given analysis frame. The phase is unimportant to third-octave analysis, thus it is discarded. Similarly, negative frequency components contain the same information as positive ones because the base signal is real. We then compute third-octave bands from the squared spectral magnitude by matrix multiplication with the frequency weights proposed in \cite{antoni2010} for $l = 2$\footnote{When implementing said method, we found that squaring the $\phi_l(p), l \geq 1$ intermediate function (p. 887) was the correct approach to meet cutoff frequencies requirements, and we believe this was the author's original intention.}. We include analysis for bands $i$ from -17 to 13, ie. center frequencies $f_i$ ranging from $20~Hz$ to $20~kHz$ with $f_0 = 1~kHz$. This means the main representation is composed of 31 values sampled every 125~ms.

@@

A Huffman coding scheme \cite{huffman1952} is then used to further reduce data dimensionality. As in most entropy coding algorithms, the efficiency of this technique depends on two important factors. A reduced amount of symbols in the dictionnary yields smaller code size on lower probability symbols. Lower data entropy, ie. lower average information content of the signal distribution, also increases performance. The first is generally obtained by applying a quantization process to the signal, whereas the second is directly linked with the probability density function (PDF) of these symbols. It is defined as $H = -\sum\limits_i p_ilog_n(p_i)$, where $p_i$ is the probability of appearance of a given symbol and $n$ the numerical base in which information is represented.

@@

In the example shown here, the former is $H_{log} = 6.24~sh$ and the latter $H_{\Delta} = 3.54~sh$. Huffman encoding is then computed with either a frame-specific symbol-code dictionnary or a constant one generated from an entire dataset. A comparison of both methods is shown in Figure~\ref{fig:dict_comp}. For most short texture frame durations, we found the local Huffman algorithm to be much faster. In fact, encoding complexity is function of the amount of dictionnary elements. If it is specific to a given data packet, then it does not need to contain every possible value. This gain seem to outweight the additional complexity induced by the tree generation algorithm. Alternatively, short texture frames or single analysis windows can be encoded using one general dictionnary to improve bitrate. This choice depends on which of the size and complexity parameters must be reduced most.\\

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.5\textwidth]{figures/pdf.eps}
	\caption{Example of estimated probability density functions of the data throughout the encoding step. (left) Unchanged output of the representation step, concentrated towards very low values. (middle) PDF "flattening" effect induced by logarithm application. Here values are mapped to the range $[0, 2^7-1]$ and rounded to perform quantization. (right) Output of the $\Delta$ compression, with desirable probabilities as the input to a Huffman algorithm.}
	\label{fig:pdf}
\end{figure}

@@

The decoding process is quite straightforward, as both Huffman and $\Delta$ compression are lossless and directly reversible. The entire coder scheme is summarized in Figure~\ref{fig:scheme}.

@@

A set of metrics is computed to assess the efficiency of the proposed coder scheme and determine the impact of the algorithm's parameters, which are as follow: the desired word size prior to Huffman encoding, set by quantization for both signal representation choices, and only for Mel bands the number of coefficients between 10 and 40. We also study the effect of time averaging analysis frames: to force a fixed amount of frames per second lesser than the average speech rate in phonemes per second is a possible way to alter intelligibility at the cost of temporal information resolution.

@@

>>>>

\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
\newcommand{\fg}[1]{\textcolor{red}{ Felix: #1}}
\newcommand{\ac}[1]{\textcolor{red}{ Arnaud: #1}}
\newcommand{\pa}[1]{\textcolor{red}{ Pierre: #1}}
\newcommand{\cl}[1]{\textcolor{red}{ Catherine: #1}}

@@

The spreading of urban areas and the growth of human population worldwide raised societal and environmental concerns. To better address those, the monitoring the acoustic environnment in urban but also rural or wilderness areas is among other monitoring efforts an important matter. Fostering on the recent development of low cost hardware acoustic sensors, we propose in this paper to consider a sensor grid approach to tackle this issue.

In this kind of approach, the crucial question is the nature of the data that is transmitted from the sensors to the processing and archival servers. To this end, we propose an efficient audio coding scheme based on third octave band spectral representation that allows 1) the estimation of standard acoustic indicators and 2) the recognition of acoustic events at state of the art performance rate. The former is useful to provide quantitative information about the acoustic environment, while the latter is useful to gather qualitative information and build perceptually motivated indicators using for example the emergence of a given sound source. The coding scheme is also demonstrated to transmit spectrally encoded data that, reverted to the time domain using state of the art techniques, is not intelligible, thus protecting the privacy of citizens.

@@

The advent of low cost acoustic sensors together with the need to better monitor and comprehend the acoustic environment of urban and wilderness areas give rise to the deployment of experimental sensor networks such as the sonyc\footnote{\url{htttp://wp.nyu.edu/sonyc}} \cite{mydlarz2017implementation} and cense\footnote{\url{htttp://cense.ifsttar.fr}} \pa{ref papier de judicael ?} \cite{} projects.

@@

In our application setting, the requirements are the following. First,  the number of sensor nodes shall be extended as needed without having to change the hardware and software architecture, \textit{i.e.} the approach shall be scalable. Second, the nodes shall be energy efficient, ideally autonomous, in order to ease the deployment of the sensors grid to the desired topology. Third, the encoding scheme used to transmit the data from the sensor to the storage and processing servers shall be designed with care in order to ensure the privacy of the citizen. Though, the data stored shall remain as rich as possible while maintaining a low bitrate for efficient transmission and storage.

@@

The latter scheme has several benefits. First, the sensor task is much simpler and can thus be implemented using low consumption processors, potentially being autonomous in terms of energy. This will ease the deployment of the network. Second, it allows researchers to gather large amount of data that can be post processed and studied further offline. Data can be re analysed following newer classification schemes or using new indicators.

However, transmitting the raw audio through the network has several disavantages, in terms of required bandwidth, storage capabilities and privacy. Thus, for transmission from the sensors to the storage unit, the data shall be encoded in a more efficient way than lossy audio encoding standards \cite{pan1995tutorial}.Indeed, the audio signal is not mandatory for the computation of the acoustic indicators and the features needed for event recognition. Also, as the data is transmitted using the network and stored, one must ensure that the intelligibility of potential speech utterances is lost during the coding process, in order to ensure the privacy of the citizens.

We thus propose in this paper a audio coding scheme specifically tailored for use in a sensor grid which has the following features. It is of low bitrate but still allows the computation of most of the standard acoustics indicators with controlable precision. As far as acoustic event detection is concerned, we report equivalent performance of several state of the art classification schemes from features computed using raw audio data and encoded data at production level rates. Finally, according to preliminary perceptual evaluation, the proposed coding scheme very strongly degrades the intelligibility, thus ensuring citizen privacy.

In order to promote reproducible research, the coder as well as the experiments needed to generate the figures is available to the community\footnote{\url{https://github.com/felixgontier/cense-coder}}. The remaining of the paper is organized as follows: Section \ref{sec:background} reviews the types of data that needs to be estimated using the transmitted data and Section \ref{sec:coder} describes the proposed encoding scheme. The experimental protocol designed to validate the proposed approach is described in Section \ref{sec:protocol} and results are discussed in Section \ref{sec:results}.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figures/censeCoder}
\vspace{-0.15in}
\caption{Alternative implementations of the sensor grid approach for the monitoring of the acoustic environnment. In A) the raw audio is transmitted, in B) only the detected events, and in C) a compressed spectral representation. Thickness of arrows roughly indicate bandwidth and thickness of boxes roughly indicates the level of computation or storage required.}
\label{fig:codingScheme}
\end{figure}

@@

The present work is constrained to allow both acoustic monitoring and sound event recognition in urban soundscapes. Here we briefly review state-of-the-art methods in both subjects in order to better motivate our design choices.

@@

The considered sensor network primarily aims at monitoring urban soundscapes, that is, continously assessing their content and impact on the population. In the literature, this is typically performed by measuring energetic acoustic indicators such as the equivalent sound pressure level $L_{eq}$ in dB SPL or its A-weighted equivalent $L_{Aeq}$ in dBA. However, while these indicators have proven to correlate well with perceptual evaluations for negative impact sound environments \cite{gozalo2015}, they are not sufficient to fully describe soundscapes \cite{rychtarikova2013}. Many other variables can be derived to better account for previously implicit properties \cite{can2016} including percentile values or time evolution. Studies have been conducted to select relevant subsets of descriptors in sound environment characterization \cite{can2015, brocolini2013, nilsson2007}.\\

@@

Eventhough the above discussed indicators provide more richness than $L_{eq}$, recent studies show that abstract statistics are still not sufficient to fully model the human perception of soundscapes. The recognition of sources of interest is thus of importance. As far as the encoding scheme is concerned, it is thus important that the encoded data allows us to compute the above cited indicators but also to allow us to recognize the sources of interest with state of the art methods.

The recognition of sources of interest from audio streams has been the subject of extensive research in the past on speech \cite{anusuya2009}, music \cite{tzanetakis2002}, and lately more complex scenes in which the current work falls. Studied classification methods are diverse, ranging from time-dependant modeling with hidden markov models \cite{ntalampiras2014} to "bag-of-frames" approaches \cite{aucouturier2007, foggia2015}. Common architectures include learning-based classifiers such as support vector machines \cite{kumar2016}, gaussian mixture models \cite{radhakrishnan2005} or neural networks \cite{salamon2017, piczak2015}. However, the selection of relevant features is still an open debate. The most used are certainly spectral \cite{khunarsal2013} or cepstral \cite{couvreur2004} representations of the signal. Among them, mel spectrograms and their cepstral-domain derivation, the mel-frequency cepstrum coefficients (MFCC), are the most recurrent. These representations effectively model the human cochlear response to sounds by grouping frequency components around critical bands in a logarithmic scale.

They may also be exploited together with features computed in other domains to better model signal properties. For instance, \cite{cai2006} adds spectral features related to harmonicity and salient frequencies, and \cite{chu2009} uses a matching pursuit (MP) algorithm to deduce time-domaid features. Another promising solution is feature engineering via unsupervised learning, which \cite{salamon2015-2} implements with a k-means clustering technique. Alternative data representations such as the scattering transform \cite{bauge2013} show good results in environmental sound classification tasks \cite{salamon2015}.\\
A more detailed review of used methods is available in \cite{chachada2013}.

@@

The technical constraints of size inherent to the studied setup make it impossible to transmit raw audio recordings. Following the considerations briefly exposed in the previous section, we conclude that third-octave bands are a reasonable choice. As mentionned, they provide advantages in both acoustic monitoring quality and reduced data volume per measurement period. The resulting representation is also close to that of mel spectrograms, being just another filterbank transform with logarithmic frequency scaling.\\

@@

The first step is to represent the signal in the frequency domain using a short-term Fourier transform (STFT). The continuously sampled audio is segmented into 125~ms frames to provide "fast" acoustic measurements. The data is zero-padded to the next power of two to improve computational efficiency. No overlap is used as it doubles both computational and memory costs and is not required for our application. A windowing function can be applied, for instance a rectangular window to ensure energy conservation in a given analysis frame. The phase is discarded as it is not meaningful for third-octave analysis. Similarly, negative frequency components contain the same information as positive ones because the input signal is real. We then compute third-octave bands from the squared spectral magnitude by matrix multiplication with the frequency weights proposed in \cite{antoni2010} for $l = 2$\footnote{When implementing said method, we found that squaring the $\phi_l(p), l \geq 1$ intermediate function (p. 887) was the correct approach to meet cutoff frequencies requirements, and we believe this was the author's original intention.}. We include analysis for bands $i$ from -17 to 13, ie. center frequencies $f_i$ ranging from $20~Hz$ to $20~kHz$ with $f_0 = 1~kHz$. The representation is thus composed of 31 values sampled every 125~ms.

@@

A Huffman coding scheme \cite{huffman1952} is then used to further reduce data rate. As in most entropy coding algorithms, the efficiency of this technique depends on two important factors. A reduced amount of symbols in the dictionnary yields smaller code size on lower probability symbols. Lower data entropy, \textit{i.e.} lower average information content of the signal distribution, also increases efficiency. The first is generally obtained by applying a quantization process to the signal, whereas the second is directly linked with the probability density function (PDF) of these symbols. It is defined as $H = -\sum\limits_i p_ilog_n(p_i)$, where $p_i$ is the probability of appearance of a given symbol and $n$ the numerical base in which information is represented.

@@

In the example shown here, the former is $H_{log} = 6.24~sh$ and the latter $H_{\Delta} = 3.54~sh$. Huffman encoding is then computed with either a frame-specific symbol-code dictionnary or a constant one generated from an entire dataset. A comparison of both methods is shown in Figure~\ref{fig:dict_comp}. For most short texture frame durations, we found the local Huffman algorithm to be much faster. In fact, encoding complexity is function of the amount of dictionnary elements. If it is specific to a given data packet, then it does not need to contain every possible value. This gain seem to outweight the additional complexity induced by the need to transmit the result of the tree generation algorithm. Alternatively, short texture frames or single analysis windows can be encoded using one general dictionnary to improve bitrate. Choosing between those two approaches is the result of a balance between requirements in terms of size and complexity.\\

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.5\textwidth]{figures/pdf.eps}
	\caption{Example of estimated probability density functions of the data throughout the encoding step. Unchanged output of the representation step (left), concentrated towards very low values. PDF "flattening" effect induced by logarithm application (middle). Here values are mapped to the range $[0, 2^7-1]$ and rounded to perform quantization. Output of the $\Delta$ compression (right) -, with desirable probabilities as the input to a Huffman algorithm.}
	\label{fig:pdf}
\end{figure}

@@

The decoding process is quite straightforward, as both Huffman and $\Delta$ compression are lossless and invertible directly. The entire coder scheme is summarized in Figure~\ref{fig:scheme}.

@@

A set of metrics is computed to assess the efficiency of the proposed coder scheme and to determine the impact of the degrees of freedom of the propoded algorithm. Those parameters are as follow: 1) the desired word size prior to Huffman encoding, set by quantization for both signal representation choices, and 2) the number of coefficients between 10 and 40 (only for Mel bands). We also study the effect of time averaging analysis frames: to force a fixed amount of frames per second lesser than the average speech rate in phonemes per second is a possible way to alter intelligibility at the cost of temporal information resolution.

///

@@

An audio encoding scheme have been described and motivated for use in a sensor grid approach to the problem of acoustic monitoring.

The proposed approach have the following benefits:
\begin{itemize}
	\item precision: the impact of the quantization step for 8 bits codewords is found to be negligible compared to sensor precision of class 2 sensors
	\item efficiency: for fast and slow modes, the bitrate is respectively about \fg{numbers} kbps for high precision
	\item privacy: at all operating modes, the reverted audio signal is found to be inintelligible.
\end{itemize}

As those constraints are to some extent orthogonal, the proposed approach is flexible enough to balance rapidly between several modes of operations by changing a few parameters in an intuitive manner.

Future work will study the usefulness of more sophisticated encoding paradigms in order to further improve bitrate efficiency and a more torough investigation on the inversion methods in order to ensure that the inintelligibility constraint is not violated.
